---
title: "Retail ananlysis"
author: "Gyaneshwar Singh"
date: "6/4/2021"
output:
  html_document:
    df_print: paged
---

```{r message=F, warning=FALSE}
# install.packages('Hmisc')
# install.packages('dplyr')
# install.packages('ggplot2')

library(dplyr)
library(ggplot2)
library(Hmisc)
library(lubridate)
```

*Setting and getting working directory*
```{r}
path = "/Users/gyaneshwarsingh/Downloads/Retail Analysis"
setwd(path)
getwd()
```

*Reading the files*

```{r}
customer_full_df <- read.csv(
  'Customer.csv',
  header = TRUE,
  stringsAsFactors = FALSE,
  na.strings = c("", "NA")
)

transactions_full_df <- read.csv(
  'Transactions.csv',
  header = TRUE,
  stringsAsFactors = FALSE,
  na.strings = c("", "NA")
)

prodcat_full_df <- read.csv(
  'prod_cat_info.csv',
  header = TRUE,
  stringsAsFactors = FALSE,
  na.strings = c("", "NA")
)

```

**1. Merge the datasets Customers, Product Hierarchy and Transactions as Customer_Final. Ensure to keep all customers who have done transactions with us and select the join type accordingly.**

**1. a) Merge using the base merge()**

```{r}
trans_cust_df1 <- merge(
  transactions_full_df,
  customer_full_df,
  by.x = "cust_id",
  by.y = "customer_Id",
  all.x = TRUE,
  all.y = FALSE
)

cust_fin1 <- merge(
  trans_cust_df1,
  prodcat_full_df,
  by.x = c("prod_cat_code", "prod_subcat_code"),
  by.y = c("prod_cat_code", "prod_sub_cat_code"),
  all.x = TRUE,
  all.y = FALSE
)

cat('Dimensions of dataframe obtained through merge(): ', dim(cust_fin1))
```

*1. b) Merge using Dplyr merge functions*

```{r}
trans_cust_df2 <- left_join(transactions_full_df,
                            customer_full_df,
                            by = c("cust_id" = "customer_Id"))

Customer_Final <- left_join(
  trans_cust_df1,
  prodcat_full_df,
  by = c("prod_cat_code" = "prod_cat_code",
         "prod_subcat_code" = "prod_sub_cat_code")
)

cat('Dimensions of dataframe obtained through dplyr(): ', dim(Customer_Final))
```

**2. Prepare a summary report for the merged data set.**

*2. a) Get the column names and their corresponding data types.*

```{r}
names(Customer_Final)
str(Customer_Final)
```

*2. b) Top/Bottom 10 observations*

```{r}
head(Customer_Final,10)
tail(Customer_Final,10)
```

*2. c) “Five-number summary” for continuous variables (min, Q1, median,Q3 and max)*

```{r}
summary(Customer_Final)
```

*2. d) Frequency tables for all the categorical variables*

```{r}
  
product_cat<- as.data.frame(table(Customer_Final$prod_cat))
names(product_cat)[1] = "Product_category"
product_subcat = as.data.frame(table(Customer_Final$prod_subcat))
names(product_subcat)[1]= "Product_SubCategories"
product_store_type<- as.data.frame(table(Customer_Final$Store_type))
names(product_store_type)[1] = "Product_store_Type"
customer_gender <- as.data.frame(table(Customer_Final$Gender))
names(customer_gender)[1] = "Gender"

```

**3. Generate histograms for all continuous variables and frequency bars for categorical variables.**

*Histograms for continuous variables*

```{r}
hist(Customer_Final$prod_subcat_code, ,main = "Histogram for Product Subcategories", xlab = "prodect subcategoires code", ylab = "total count", border = "blue", col = "purple",)

hist(Customer_Final$prod_cat_code, main = "histogram for product categories", xlab = "product Categories code", ylab = "total count", border = "black", col = "orange")

hist(Customer_Final$city_code, main = "Histogram Based on City Code", xlab = "City Code", ylab = "total Count", border = "black", col = "pink")
```

*Frequency bars for continuous variables*

```{r}
Product_subcat_code <- as.data.frame(table(Customer_Final$prod_subcat_code))
 names(Product_subcat_code)[1] = "Product Subcategory Code"
Product_cat_code <- as.data.frame(table(Customer_Final$prod_cat_code))
names(Product_cat_code)[1] = "Product Categorey Code"
Product_city_code <- as.data.frame(table(Customer_Final$city_code))
names(Product_city_code)[1] = "product City code"
```

**4. Calculate the following information using the merged dataset :**

*4 a) Calculate Time period of the available transaction data.*

```{r}

```

*4 b) Calculate count of transactions where the total amount of transaction was negative*

```{r}
length(Customer_Final$total_amt < 0)
```

**5. Analyze which product categories are more popular among females vs male customers.**

*Most popular categories among males*

```{r}

d <- data.frame(filter(Customer_Final, Gender == "M"))
mytable <- as.data.frame(table(d$prod_cat))
names(mytable)[1] = "Popular Catergies"
arrange(mytable, desc(mytable$Freq))

```

*Most popular categories among Females*

```{r}
df <- data.frame(filter(Customer_Final, Gender == "F"))
mytable2 <- as.data.frame(table(df$prod_cat))
names(mytable2)[1] = "Popular Catergies"
arrange(mytable2, desc(mytable2$Freq))


```

*The top 3 categories for both males and females are:*

* Books

* Electronics

* Home and kitchen

**6. Which City code has the maximum customers and what was the percentage of customers from that city?**

*Percentage of customers belonging to different city codes sorted in descending order:*

```{r}
customer_percentage <- (Product_city_code$Freq / sum(Product_city_code$Freq))*100
product_city_percentage<- cbind(Product_city_code,customer_percentage)
arrange(product_city_percentage, desc(product_city_percentage$customer_percentage))
```

The highest percentage of customers come from city code 4 (10.51%)

**7. Which store type sells the maximum products by value and by quantity?**

*Products sold by value:*

```{r}
aa <- as.data.frame(table(Customer_Final$Qty))
names(aa)[1]= "Quantity"
arrange(aa, desc(Quantity))
```


**8. What was the total amount earned from the "Electronics" and "Clothing"categories from Flagship Stores?**

```{r}
ddf <- data.frame(filter(Customer_Final, (prod_cat == "Electronics" | prod_cat == "Clothing") & Store_type == "Flagship store" ))
sum(ddf$total_amt)

```

**9. What was the total amount earned from "Male" customers under the "Electronics" category?**

```{r}
male_sum <- data.frame(filter(Customer_Final, prod_cat =="Electronics" & Gender =="M"))
sum(male_sum$total_amt)
```

**10. How many customers have more than 10 unique transactions, after removing all transactions which have any negative amounts?**

```{r}
sd<- data.frame(filter(Customer_Final,total_amt > 0 ))
trans_count <- group_by(sd,cust_id)
y<- summarise(trans_count, total_transaction_id = n_distinct(transaction_id))
y[y$total_transaction_id >10,]

```
**11a. What was the total amount spent for “Electronics” and “Books”product categories?**
```{r}
b_sum <- data.frame(filter(Customer_Final, prod_cat =="Electronics" | prod_cat == "Books"))
sum(b_sum$total_amt)
```

**11b. What was the total amount spent by these customers between 1stJan, 2014 to 1st Mar, 2014?**
```{r}
#d.sum <- data_frame(filter(Customer_Final,tran_date > as.Date('01/01/2014') & tran_date < as.Date('03/01/2014')))
#sum(d.sum$total_amt)
```


